---
title: "Extract SST data from OISST and CRW .nc files"
author: "Ana M. Palacio-Castro"
date: "September 10, 2018"
output:
  html_document:
    toc: true
    theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, results="hide",)
```

```{r, message=FALSE}
  library(raster)
  library(parallel)
```
 
# 1. OISST Data 

This chunk of the script use the .nc files obtained from ftp://ftp.cdc.noaa.gov/Datasets/noaa.oisst.v2.highres/ (data from 1981-09-01 to 2016-12-31 in the "DHW/Daily_OISST_v2" folder) to extract the SST data for the closest pixel to Uva Island reef (OISST coordinates = 7.875, 278.125)

### Extract OISST temperature data based on coordinates

```{r Extract_OISST_data, cache=TRUE}

   # read all the .nc files 
    flist<-list.files('Daily_OISST_v2',pattern='*.nc',full.names=TRUE)
      
  # brick them 
    ras<-lapply(flist, stack)
    
  # Uva Island Reef coordinates   
    lon.pts <-(278.125) 
    lat.pts <- (7.875)
    extract.pts <- cbind(lon.pts,lat.pts)

  ## Extract data by coordinates 
    ext <- lapply(ras,raster::extract,extract.pts)
    #head(ext)
    OISST <- unlist(ext, recursive = FALSE)
    #head(SST)
```

### Bind OISST temperature data with dates

```{r Bind_SST_GPS_Date}
  
# Create a date list 
    Dates<-seq.Date(as.Date("1981-09-01"), as.Date("2016-12-31"), by = "day")
    summary(Dates)
    Date<-as.character(Dates)
    #summary(Date)
    #head(Date)
      
# Bind GPS points and Dates    Date.location <- merge(pts, Date)
    OISST.data<-data.frame(Date, OISST)
    OISST.data$Date<-as.Date(OISST.data$Date)
    # head(OISST.data)  
    # summary(OISST.data)
    write.csv(OISST.data, "Uva_Daily_OISST_1982-2016.csv", row.names=FALSE)
```

# 2. CRW Data (CoralTemp) 

This chunk of the script use the .nc files obtained from ftp://ftp.star.nesdis.noaa.gov/pub/sod/mecb/crw/data/coraltemp/v1.0/nc/ (data from 1985 to 2016 in the "Daily_CoralTemp_3.1" folder) to extract the SST data for the closest pixel to Uva Island reef CRW_5km coordinates = 7.825003, -81.77499)

```{r}
rm(list = ls())  # Remove all objects from memory
```

### Extract CRW 5km temperature data based on coordinates

```{r Extract_OISST_data, cache=TRUE}

   # read all the .nc files 
    flist<-list.files('Daily_CoralTemp_3.1',pattern='*.nc',full.names=TRUE)
      
  # brick them 
    ras<-lapply(flist, stack)
    
  # Uva Island Reef coordinates   
    lon.pts <-(-81.77499)
    lat.pts <- (7.825003)
    extract.pts <- cbind(lon.pts,lat.pts)

  ## Extract data by coordinates 
    ext <- lapply(ras,raster::extract,extract.pts)
    #head(ext)
    CRWSST <- unlist(ext, recursive = FALSE)
    #head(CRWSST)
    #summary(CRWSST)
```

### Bind CRW 5km temperature data with dates

```{r Bind_SST_GPS_Date}
  
# Create a date list 
    Dates<-seq.Date(as.Date("1985-01-01"), as.Date("2016-12-31"), by = "day")
    Date<-as.character(Dates)
    #head(Date)
      
# Bind GPS points and Dates    Date.location <- merge(pts, Date)
    CRWSST.data<-data.frame(Date, CRWSST)
    CRWSST.data$Date<-as.Date(CRWSST.data$Date)
    # head(CRWSST.data)  
    # summary(CRWSST.data)
    write.csv(CRWSST.data, "Uva_Daily_CRW_5km_1985-2016.csv", row.names=FALSE)

```

### Extract CRW 5km MMM from climatology

This chunk of the script use the .nc file obtained from ftp://ftp.star.nesdis.noaa.gov/pub/sod/mecb/crw/data/5km/v3.1/climatology/nc/ct5km_climatology_v3.1.nc cointaining the climatology calculated from the 1985-2012 preriod to extract the MMM temperature for the closest pixel to Uva Island reef CRW_5km coordinates = 7.825003, -81.77499)

```{r Extract_OISST_data, cache=TRUE}

  # Read climatology file
  Climatology_MMM <- "ct5km_climatology_v3.1.nc"
  MMM.data <- brick(Climatology_MMM) 
  #MMM.data
  
  # Uva Island Reef   
   lon.pts <-(-81.77499) 
   lat.pts <- (7.825003)
   extract.pts <- cbind(lon.pts,lat.pts)
  
  #MMM <- raster::extract(MMM.data, extract.pts,method="bilinear")
  MMM <- raster::extract(MMM.data, extract.pts,method="simple")
  write.csv(MMM, "Uva_CRW_MMM_1985-2012.csv", row.names=FALSE)
```

# 3. CRW DHW 

This chunk of the script use the .nc files obtained from ftp://ftp.star.nesdis.noaa.gov/pub/sod/mecb/crw/data/5km/v3.1/nc/v1.0/daily/dhw/ (data from 1985 to 2016 in the "Daily_CWR_DHW_3.1" folder) to extract the already calculated DHW data for the closest pixel to Uva Island reef CRW_5km coordinates = 7.825003, -81.77499)

```{r}
rm(list = ls())  # Remove all objects from memory
```

### Extract CRW 5km DHW data based on coordinates

```{r, cache=TRUE}

  #  # read all the .nc files 
  #   flist<-list.files('Daily_CWR_DHW_3.1',pattern='*.nc',full.names=TRUE)
  #     
  # # brick them 
  #   ras<-lapply(flist, stack)
  #   
  # # Uva Island Reef coordinates   
  #   lon.pts <-(-81.77499)
  #   lat.pts <- (7.825003)
  #   extract.pts <- cbind(lon.pts,lat.pts)
  # 
  # ## Extract data by coordinates 
  #   ext <- lapply(ras,raster::extract,extract.pts)
  #   #head(ext)
  #   DHW <- unlist(ext, recursive = FALSE)
  #   #head(SST)
  #   #summary(DHW)
```

### Bind CRW 5km DHW data with dates

```{r}
  
# # Create a date list 
#     Dates<-seq.Date(as.Date("1985-03-25"), as.Date("2016-12-31"), by = "day")
#     Date<-as.character(Dates)
#     #head(Date)
#       
# # Bind GPS points and Dates    Date.location <- merge(pts, Date)
#     DHW.data<-data.frame(Date, DHW)
#     DHW.data$Date<-as.Date(DHW.data$Date)
#     # head(DHW.data)  
#     # summary(DHW.data)
#     write.csv(DHW.data, "Uva_Daily_DHW_5km_1985-2016.csv", row.names=FALSE)
```
